---
title: "lab3b"
author: "Marlee_Galloway"
date: "9/7/2020"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## load the tidyverse packages, incl. dplyr
```{r}
library("tidyverse")
```

```{r}
surveys <- read_csv("data/portal_data_joined.csv")
```
```{r}
## inspect the data
str(surveys)
```

###Selecting Colums and Filtering Rows
```{r}
select(surveys, plot_id, species_id, weight)
```
```{r}
select(surveys, -record_id, -species_id)
```
```{r}
filter(surveys, year == 1995)
```

###Pipes: to select and filter at the same time
#Intermediate steps:
```{r}
surveys2 <- filter(surveys, weight < 5)
surveys_sml <- select(surveys2, species_id, sex, weight)
```
#Nested functions:
```{r}
surveys_sml <- select(filter(surveys, weight < 5), species_id, sex, weight)
```
#Pipes:
```{r}
surveys %>%
  filter(weight < 5) %>%
  select(species_id, sex, weight)
```
#Create new object with smaller data set:
```{r}
surveys_sml <- surveys %>%
  filter(weight < 5) %>%
  select(species_id, sex, weight)

surveys_sml
```

###CHALLENGE Q1

#Using pipes, subset the surveys data to include animals collected before 1995 and retain only the columns year, sex, and weight.

##Ans: 
```{r}
surveys %>%
    filter(year < 1995) %>%
    select(year, sex, weight)
```

###Mutate
#new column of weight in kg:
```{r}
surveys %>%
  mutate(weight_kg = weight / 1000)
```
#second new column:
```{r}
surveys %>%
  mutate(weight_kg = weight / 1000,
         weight_lb = weight_kg * 2.2)
```
#using a pipe to view the head of the data:
```{r}
surveys %>%
  mutate(weight_kg = weight / 1000) %>%
  head()
```
#to remove NA's insert a filter in the chain:
```{r}
surveys %>%
  filter(!is.na(weight)) %>%
  mutate(weight_kg = weight / 1000) %>%
  head()
```

###CHALLENGE Q2
#Create a new data frame from the surveys data that meets the following criteria: contains only the species_id column and a new column called hindfoot_cm containing the hindfoot_length values converted to centimeters. In this hindfoot_cm column, there are no NAs and all values are less than 3.

##Ans:
```{r}
surveys_hindfoot_cm <- surveys %>%
    filter(!is.na(hindfoot_length)) %>%
    mutate(hindfoot_cm = hindfoot_length / 10) %>%
    filter(hindfoot_cm < 3) %>%
    select(species_id, hindfoot_cm)
```

###Split-apply-combine data analysis and the summarize() function
#Compute the mean weight by sex:
```{r}
surveys %>%
  group_by(sex) %>%
  summarize(mean_weight = mean(weight, na.rm = TRUE))
```
#Remove missing values for weight before calc. summary stats
```{r}
surveys %>%
  filter(!is.na(weight)) %>%
  group_by(sex, species_id) %>%
  summarize(mean_weight = mean(weight))
```
#To display more data using print()
```{r}
surveys %>%
  filter(!is.na(weight)) %>%
  group_by(sex, species_id) %>%
  summarize(mean_weight = mean(weight)) %>%
  print(n = 15)
```
#summarize multiple variables at the same time:
```{r}
surveys %>%
  filter(!is.na(weight)) %>%
  group_by(sex, species_id) %>%
  summarize(mean_weight = mean(weight),
            min_weight = min(weight))
```
#Rearrange the result to inspect values:
```{r}
surveys %>%
  filter(!is.na(weight)) %>%
  group_by(sex, species_id) %>%
  summarize(mean_weight = mean(weight),
            min_weight = min(weight)) %>%
  arrange(min_weight)
```
#To sort results in decreasing order of mean weight:
```{r}
surveys %>%
  filter(!is.na(weight)) %>%
  group_by(sex, species_id) %>%
  summarize(mean_weight = mean(weight),
            min_weight = min(weight)) %>%
  arrange(desc(mean_weight))
```
##Counting
#Number of rows of data for each sex:
```{r}
surveys %>%
    count(sex)
```
#The count() fx is shorthand for:
```{r}
surveys %>%
    group_by(sex) %>%
    summarise(count = n())

```
#Sort 
```{r}
surveys %>%
    count(sex, sort = TRUE) 

```
#To count a combination of factors:
```{r}
surveys %>%
  count(sex, species)
```
#To sort table according to a number of criteria:
```{r}
surveys %>%
  count(sex, species) %>%
  arrange(species, desc(n))
```

###CHALLENGE Q3
#1. How many animals were caught in each plot_type surveyed?
```{r}
surveys %>%
    count(plot_type) 
```

#2. Use group_by() and summarize() to find the mean, min, and max hindfoot length for each species (using species_id). Also add the number of observations (hint: see ?n).
```{r}
surveys %>%
    filter(!is.na(hindfoot_length)) %>%
    group_by(species_id) %>%
    summarize(
        mean_hindfoot_length = mean(hindfoot_length),
        min_hindfoot_length = min(hindfoot_length),
        max_hindfoot_length = max(hindfoot_length),
        n = n()
    )
```

#3. What was the heaviest animal measured in each year? Return the columns year, genus, species_id, and weight.
```{r}
surveys %>%
    filter(!is.na(weight)) %>%
    group_by(year) %>%
    filter(weight == max(weight)) %>%
    select(year, genus, species, weight) %>%
    arrange(year)
```

###Reshaping with gather and spread
#Spreading: spread()
```{r}
surveys_gw <- surveys %>%
  filter(!is.na(weight)) %>%
  group_by(plot_id, genus) %>%
  summarize(mean_weight = mean(weight))
```
```{r}
str(surveys_gw)
```
#Use spread() to key on genus with values from mean_weight
```{r}
surveys_spread <- surveys_gw %>%
  spread(key = genus, value = mean_weight)

str(surveys_spread)
```
#Fill in missing values:
```{r}
surveys_gw %>%
  spread(genus, mean_weight, fill = 0) %>%
  head()

```
#Gathering: gather()
```{r}
surveys_gather <- surveys_spread %>%
  gather(key = "genus", value = "mean_weight", -plot_id)

str(surveys_gather)
```
#Specification of columns:
```{r}
surveys_spread %>%
  gather(key = "genus", value = "mean_weight", Baiomys:Spermophilus) %>%
  head()

```

###CHALLENGE Q4
#1. Spread the surveys data frame with year as columns, plot_id as rows, and the number of genera per plot as the values. You will need to summarize before reshaping, and use the function n_distinct() to get the number of unique genera within a particular chunk of data. It’s a powerful function! See ?n_distinct for more.
```{r}
surveys_spread_genera <- surveys %>%
  group_by(plot_id, year) %>%
  summarize(n_genera = n_distinct(genus)) %>%
  spread(year, n_genera)
```
```{r}
head(surveys_spread_genera)
```
#2. Now take that data frame and gather() it again, so each row is a unique plot_id by year combination.
```{r}
surveys_spread_genera %>%
  gather("year", "n_genera", -plot_id)
```
#3. The surveys data set has two measurement columns: hindfoot_length and weight. This makes it difficult to do things like look at the relationship between mean values of each measurement per year in different plot types. Let’s walk through a common solution for this type of problem. First, use gather() to create a dataset where we have a key column called measurement and a value column that takes on the value of either hindfoot_length or weight. Hint: You’ll need to specify which columns are being gathered.
```{r}
surveys_long <- surveys %>%
  gather("measurement", "value", hindfoot_length, weight)
```
#4. With this new data set, calculate the average of each measurement in each year for each different plot_type. Then spread() them into a data set with a column for hindfoot_length and weight. Hint: You only need to specify the key and value columns for spread().
```{r}
surveys_long %>%
  group_by(year, measurement, plot_type) %>%
  summarize(mean_value = mean(value, na.rm=TRUE)) %>%
  spread(measurement, mean_value)
```

###Exporting data
#Removing observations of animals for which weight and hindfoot_length are missing, or the sex has not been determined:
```{r}
   surveys_complete <- surveys %>%
  filter(!is.na(weight),           # remove missing weight
         !is.na(hindfoot_length),  # remove missing hindfoot_length
         !is.na(sex))                # remove missing sex            
```
#Remove observations of rare species:
```{r}
## Extract the most common species_id
species_counts <- surveys_complete %>%
    count(species_id) %>% 
    filter(n >= 50)

## Only keep the most common species
surveys_complete <- surveys_complete %>%
  filter(species_id %in% species_counts$species_id)
```
```{r}
dim(surveys_complete)
```
#Save as csv file in data folder:
```{r}
write_csv(surveys_complete, path = "data/surveys_complete.csv")
```

