---
title: "lab3A"
author: "Marlee_Galloway"
date: "9/7/2020"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
```

```{r eval=FALSE}
 download.file(url="https://ndownloader.figshare.com/files/2292169",
              destfile = "data/portal_data_joined.csv")
```

```{r}
surveys <- read_csv("data/portal_data_joined.csv")
```

```{r}
surveys
```

```{r}
head(surveys)
```

```{r}
str(surveys)
```

### Inspecting data.frame objects

#Size:
#dim(surveys) - returns a vector with the number of rows in the first element, and the number of columns as the second element (the dimensions of the object)
#nrow(surveys) - returns the number of rows
#ncol(surveys) - returns the number of columns

#Content:
#head(surveys) - shows the first 6 rows
#tail(surveys) - shows the last 6 rows

#Names:
#names(surveys) - returns the column names (synonym of colnames() for data.frame objects)
#rownames(surveys) - returns the row names

#Summary:
#str(surveys) - structure of the object and information about the class, length and content of each column
#summary(surveys) - summary statistics for each column

### CHALLENGE Q1
#Based on the output of str(surveys), can you answer the following questions?
#What is the class of the object surveys?
#How many rows and how many columns are in this object?
#How many species have been recorded during these surveys?

##Ans:
## * class: data frame
## * how many rows: 34786,  how many columns: 13
## * how many species: 48

###Indexing and Subsetting Data Frames
#Extract specific data:
```{r}
# first element in the first column of the data frame (as a vector)
surveys[1, 1]   
# first element in the 6th column (as a vector)
surveys[1, 6]   
# first column of the data frame (as a vector)
surveys[, 1]    
# first column of the data frame (as a data.frame)
surveys[1]      
# first three elements in the 7th column (as a vector)
surveys[1:3, 7] 
# the 3rd row of the data frame (as a data.frame)
surveys[3, ]    
# equivalent to head_surveys <- head(surveys)
head_surveys <- surveys[1:6, ] 
```
```{r}
surveys[1:10]
```
```{r}
surveys[10:1]
```
#Subsetting data frames by calling their column names
```{r}
#surveys["species_id"]       # Result is a data.frame
#surveys[, "species_id"]     # Result is a vector
#surveys[["species_id"]]     # Result is a vector
#surveys$species_id          # Result is a vector      
```

###CHALLENGE Q2
#Create a data.frame (surveys_200) containing only the data in row 200 of the surveys dataset.
#Notice how nrow() gave you the number of rows in a data.frame?
#*Use that number to pull out just that last row in the data frame.
#*Compare that with what you see as the last row using tail() to make sure it’s meeting expectations.
#*Pull out that last row using nrow() instead of the row number.
#*Create a new data frame (surveys_last) from that last row.
#Use nrow() to extract the row that is in the middle of the data frame. Store the content of this row in an object named surveys_middle.
#Combine nrow() with the - notation above to reproduce the behavior of head(surveys), keeping just the first through 6th rows of the surveys dataset.

##Ans:
```{r}
## 1.
surveys_200 <- surveys[200, ]
## 2.
# Saving `n_rows` to improve readability and reduce duplication
n_rows <- nrow(surveys)
surveys_last <- surveys[n_rows, ]
## 3.
surveys_middle <- surveys[n_rows / 2, ]
## 4.
surveys_head <- surveys[-(7:n_rows), ]
```

###Factors
```{r}
sex <- factor(c("male", "female", "female", "male"))
```
```{r}
levels(sex)
nlevels(sex)
```
```{r}
sex # current order
sex <- factor(sex, levels = c("male", "female"))
sex # after re-ordering
```

###Converting Factors

#converting a factor to a character vector:
```{r}
as.character(sex)
```
#conversion from factor to numeric:
```{r}
year_fct <- factor(c(1990, 1983, 1977, 1998, 1990))
as.numeric(year_fct)               # Wrong! And there is no warning...
as.numeric(as.character(year_fct)) # Works...
as.numeric(levels(year_fct))[year_fct]    # The recommended way.
```

###Renaming Factors
## bar plot of the number of females and males captured during the experiment:
```{r}
plot(as.factor(surveys$sex))
```
```{r}
sex <- factor(surveys$sex)
head(sex)
```
```{r}
levels(sex)
```
```{r}
levels(sex)[1] <- "undetermined"
levels(sex)
```
```{r}
head(sex)
```

###CHALLENGE Q3
#Rename “F” and “M” to “female” and “male” respectively.
#Now that we have renamed the factor level to “undetermined”, can you recreate the barplot such that “undetermined” is last (after “male”)?

##Ans:
```{r}
levels(sex)[2:3] <- c("female", "male")
sex <- factor(sex, levels = c("female", "male", "undetermined"))
plot(sex)
```

###Using stringsAsFactors = FALSE
```{r}
## Compare the difference between our data read as `factor` vs `character`.
surveys <- read.csv("data/portal_data_joined.csv", stringsAsFactors = TRUE)
str(surveys)
surveys <- read.csv("data/portal_data_joined.csv", stringsAsFactors = FALSE)
str(surveys)
## Convert the column "plot_type" into a factor
surveys$plot_type <- factor(surveys$plot_type)

```

###CHALLENGE Q4
#1 We have seen how data frames are created when using read.csv(), but they can also be created by hand with the data.frame() function. There are a few mistakes in this hand-crafted data.frame. Can you spot and fix them? Don’t hesitate to experiment!

#animal_data <- data.frame(
#          animal = c(dog, cat, sea cucumber, sea urchin),
#          feel = c("furry", "squishy", "spiny"),
#          weight = c(45, 8 1.1, 0.8)
#          )

#2 Can you predict the class for each of the columns in the following example? Check your guesses using str(country_climate):
#Are they what you expected? Why? Why not?
#What would have been different if we had added stringsAsFactors = FALSE when creating the data frame?
#What would you need to change to ensure that each column had the accurate data type?

#country_climate <- data.frame(
#       country = c("Canada", "Panama", "South Africa", "Australia"),
#       climate = c("cold", "hot", "temperate", "hot/temperate"),
#       temperature = c(10, 30, 18, "15"),
#       northern_hemisphere = c(TRUE, TRUE, FALSE, "FALSE"),
#       has_kangaroo = c(FALSE, FALSE, FALSE, 1)
#       )

##Ans: 
#missing quotations around the names of the animals
#missing one entry in the feel column (probably for one of the furry animals)
#missing one comma in the weight column
#country, climate, temperature, and northern_hemisphere are factors; has_kangaroo is numeric
#using stringsAsFactors = FALSE would have made character vectors instead of factors
#removing the quotes in temperature and northern_hemisphere and replacing 1 by TRUE in the has_kangaroo column would give what was probably intended

###Formatting Dates
```{r}
str(surveys)
```
```{r}
library("lubridate")
```
```{r}
my_date <- ymd("2015-01-01")
str(my_date)
```
```{r}
my_date <- ymd(paste("2015", "1", "1", sep = "-")) 
str(my_date)
```
#paste(surveys$year, surveys$month, surveys$day, sep = "-")
#ymd(paste(surveys$year, surveys$month, surveys$day, sep = "-"))
```{r}
surveys$date <- ymd(paste(surveys$year, surveys$month, surveys$day, sep = "-"))
```
```{r}
str(surveys) # notice the new column, with 'date' as the class
```
```{r}
summary(surveys$date)
```
```{r}
missing_dates <- surveys[is.na(surveys$date), c("year", "month", "day")]

head(missing_dates)
```


