---
title: "Lab6"
author: "Marlee_Galloway"
date: "9/29/2020"
output: html_document
bibliography: data/references.bib 
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

###Building Maps

```{r}
library(tidyverse)
library(maps)
library(mapdata)
library(lubridate)
library(viridis)
library(wesanderson)
```

```{r}
daily_report <- read_csv(url("https://raw.githubusercontent.com/CSSEGISandData/COVID-19/master/csse_covid_19_data/csse_covid_19_daily_reports/04-02-2020.csv")) %>% 
  rename(Long = "Long_") 
    
ggplot(daily_report, aes(x = Long, y = Lat, size = Confirmed/1000)) +
    borders("world", colour = NA, fill = "grey90") +
    theme_bw() +
    geom_point(shape = 21, color='purple', fill='purple', alpha = 0.5) +
    labs(title = 'World COVID-19 Confirmed cases',x = '', y = '',
        size="Cases (x1000))") +
    theme(legend.position = "right") +
    coord_fixed(ratio=1.5)
```

##To zoom in on US states (Alaska, Hawaii, US territories, and states with lat and long of 0 have to be filtered):

```{r}
daily_report <-   read_csv(url("https://raw.githubusercontent.com/CSSEGISandData/COVID-19/master/csse_covid_19_data/csse_covid_19_daily_reports/04-05-2020.csv")) %>% 
  rename(Long = "Long_") %>% 
  filter(Country_Region == "US") %>% 
  filter (!Province_State %in% c("Alaska","Hawaii", "American Samoa",
                  "Puerto Rico","Northern Mariana Islands", 
                  "Virgin Islands", "Recovered", "Guam", "Grand Princess",
                  "District of Columbia", "Diamond Princess")) %>% 
  filter(Lat > 0)
ggplot(daily_report, aes(x = Long, y = Lat, size = Confirmed)) +
    borders("state", colour = "black", fill = "grey90") +
    theme_bw() +
    geom_point(shape = 21, color='purple', fill='purple', alpha = 0.5) +
    labs(title = 'COVID-19 Confirmed Cases in the US', x = '', y = '',
        size="Cases") +
    theme(legend.position = "right") +
    coord_fixed(ratio=1.5)
```

##More intricate example:

```{r}
mybreaks <- c(1, 100, 1000, 10000, 10000)
ggplot(daily_report, aes(x = Long, y = Lat, size = Confirmed)) +
    borders("state", colour = "white", fill = "grey90") +
    geom_point(aes(x=Long, y=Lat, size=Confirmed, color=Confirmed),stroke=F, alpha=0.7) +
    scale_size_continuous(name="Cases", trans="log", range=c(1,7), 
                        breaks=mybreaks, labels = c("1-99",
                        "100-999", "1,000-9,999", "10,000-99,999", "50,000+")) +
    scale_color_viridis_c(option="viridis",name="Cases",
                        trans="log", breaks=mybreaks, labels = c("1-99",
                        "100-999", "1,000-9,999", "10,000-99,999", "50,000+"))  +
# Cleaning up the graph
  
  theme_void() + 
    guides( colour = guide_legend()) +
    labs(title = "Anisa Dhana's lagout for COVID-19 Confirmed Cases in the US'") +
    theme(
      legend.position = "bottom",
      text = element_text(color = "#22211d"),
      plot.background = element_rect(fill = "#ffffff", color = NA), 
      panel.background = element_rect(fill = "#ffffff", color = NA), 
      legend.background = element_rect(fill = "#ffffff", color = NA)
    ) +
    coord_fixed(ratio=1.5)
```

##Use the ggplot function borders() to define the areas on a map

###Mapping Data to Shapes

```{r}
daily_report <- read_csv(url("https://raw.githubusercontent.com/CSSEGISandData/COVID-19/master/csse_covid_19_data/csse_covid_19_daily_reports/04-02-2020.csv")) %>% 
  rename(Long = "Long_") %>% 
  filter(Country_Region == "US") %>% 
  group_by(Province_State) %>% 
  summarize(Confirmed = sum(Confirmed)) %>% 
  mutate(Province_State = tolower(Province_State))
# load the US map data
us <- map_data("state")
# We need to join the us map data with our daily report to make one data frame/tibble
state_join <- left_join(us, daily_report, by = c("region" = "Province_State"))
# plot state map
```

##Using R color palettes:

#Using Wes Anderson:

```{r}
# plot state map
ggplot(data = us, mapping = aes(x = long, y = lat, group = group)) + 
  coord_fixed(1.3) + 
# Add data layer
  geom_polygon(data = state_join, aes(fill = Confirmed), color = "black") +
  scale_fill_gradientn(colours = 
                         wes_palette("Zissou1", 100, type = "continuous"),
                         trans = "log10") +
  labs(title = "COVID-19 Confirmed Cases in the US'")
```

#Looking at counties using RColorBrewer:

```{r}
library(RColorBrewer)
# To display only colorblind-friendly brewer palettes, specify the option colorblindFriendly = TRUE as follow:
# display.brewer.all(colorblindFriendly = TRUE)
# Get and format the covid report data
report_03_27_2020 <-   read_csv(url("https://raw.githubusercontent.com/CSSEGISandData/COVID-19/master/csse_covid_19_data/csse_covid_19_daily_reports/04-02-2020.csv")) %>% 
  rename(Long = "Long_") %>% 
  unite(Key, Admin2, Province_State, sep = ".") %>% 
  group_by(Key) %>% 
  summarize(Confirmed = sum(Confirmed)) %>% 
  mutate(Key = tolower(Key))
# dim(report_03_27_2020)
# get and format the map data
us <- map_data("state")
counties <- map_data("county") %>% 
  unite(Key, subregion, region, sep = ".", remove = FALSE)
# Join the 2 tibbles
state_join <- left_join(counties, report_03_27_2020, by = c("Key"))
# sum(is.na(state_join$Confirmed))
ggplot(data = us, mapping = aes(x = long, y = lat, group = group)) + 
  coord_fixed(1.3) + 
  # Add data layer
  borders("state", colour = "black") +
  geom_polygon(data = state_join, aes(fill = Confirmed)) +
  scale_fill_gradientn(colors = brewer.pal(n = 5, name = "PuRd"),
                       breaks = c(1, 10, 100, 1000, 10000, 100000),
                       trans = "log10", na.value = "White") +
  ggtitle("Number of Confirmed Cases by US County") +
  theme_bw() 
```

#Looking at just Massachusetts:

```{r}
daily_report <-   read_csv(url("https://raw.githubusercontent.com/CSSEGISandData/COVID-19/master/csse_covid_19_data/csse_covid_19_daily_reports/04-02-2020.csv")) %>% 
  rename(Long = "Long_") %>% 
  filter(Province_State == "Massachusetts") %>% 
  group_by(Admin2) %>% 
  summarize(Confirmed = sum(Confirmed)) %>% 
  mutate(Admin2 = tolower(Admin2))
us <- map_data("state")
ma_us <- subset(us, region == "massachusetts")
counties <- map_data("county")
ma_county <- subset(counties, region == "massachusetts")
state_join <- left_join(ma_county, daily_report, by = c("subregion" = "Admin2")) 
# plot state map
ggplot(data = ma_county, mapping = aes(x = long, y = lat, group = group)) + 
  coord_fixed(1.3) + 
# Add data layer
  geom_polygon(data = state_join, aes(fill = Confirmed), color = "white") +
    scale_fill_gradientn(colors = brewer.pal(n = 5, name = "BuGn"),
                         trans = "log10") +
  labs(title = "COVID-19 Confirmed Cases in Massachusetts'")
```

```{r}
daily_report
```

###Interactive Graphs

##Use plotly to make interactive graphs with the map

```{r}
library(plotly)
ggplotly(
  ggplot(data = ma_county, mapping = aes(x = long, y = lat, group = group)) + 
  coord_fixed(1.3) + 
# Add data layer
  geom_polygon(data = state_join, aes(fill = Confirmed), color = "black") +
    scale_fill_gradientn(colours = 
                         wes_palette("Zissou1", 100, type = "continuous")) +
  ggtitle("COVID-19 Cases in MA") +
# Cleaning up the graph
  labs(x=NULL, y=NULL) +
  theme(panel.border = element_blank()) +
  theme(panel.background = element_blank()) +
  theme(axis.ticks = element_blank()) +
  theme(axis.text = element_blank())
)
```

##Example with the world map:

```{r}
# Read in the daily report
daily_report <- read_csv(url("https://raw.githubusercontent.com/CSSEGISandData/COVID-19/master/csse_covid_19_data/csse_covid_19_daily_reports/09-26-2020.csv")) %>% 
  rename(Long = "Long_") %>% 
  group_by(Country_Region) %>% 
  summarize(Confirmed = sum(Confirmed), Deaths = sum(Deaths))

# Read in the world map data
world <- as_tibble(map_data("world"))

# Check to see if there are differences in the naming of countries
setdiff(world$region, daily_report$Country_Region) 
```

```{r}
# Many of these countries are considered states or territories in the JHU covid reports,
# but let's fix a few of them

world <- as_tibble(map_data("world")) %>% 
 mutate(region = str_replace_all(region, c("USA" = "US", "Czech Republic" = "Czechia",  
        "Ivory Coast" = "Cote d'Ivoire", "Democratic Republic of the Congo" = "Congo (Kinshasa)", 
        "Republic of Congo" = "Congo (Brazzaville)")))

# Join the covid report with the map data
country_join <- left_join(world, daily_report, by = c("region" = "Country_Region"))

# Create the graph
ggplotly(
ggplot(data = world, mapping = aes(x = long, y = lat, text = region, group = group)) + 
  coord_fixed(1.3) + 
# Add data layer
  geom_polygon(data = country_join, aes(fill = Deaths), color = "black") +
  scale_fill_gradientn(colours = 
                         wes_palette("Zissou1", 100, type = "continuous")) +
  labs(title = "COVID-19 Deaths'")
)
```

###EXERCISES

#1. 
```{r}
daily_report_9_26_2020 <- read_csv(url("https://raw.githubusercontent.com/CSSEGISandData/COVID-19/master/csse_covid_19_data/csse_covid_19_daily_reports/09-26-2020.csv")) %>% 
  rename(Long = "Long_") %>% 
  filter(Lat != "NA") %>% 
  filter(Long != "NA") %>% 
  group_by(Country_Region) %>% 
  summarize(Deaths = sum(Deaths), Lat = median(Lat), Long = median(Long))

ggplot(daily_report_9_26_2020, mapping = aes(x = Long, y = Lat, size = Deaths/1000)) +
    borders("world", colour = NA, fill = "grey90") +
    theme_bw() +
    geom_point(shape = 21, color='purple', fill='purple', alpha = 0.5) +
    labs(title = 'World COVID-19 Deaths',x = '', y = '',
        size="Cases (x1000))") +
    theme(legend.position = "right") +
    coord_fixed(ratio=1.5)
```

#2. 
```{r}
daily_report_9_26_2020_US <-   read_csv(url("https://raw.githubusercontent.com/CSSEGISandData/COVID-19/master/csse_covid_19_data/csse_covid_19_daily_reports/09-26-2020.csv")) %>% 
  rename(Long = "Long_") %>% 
  filter(Country_Region == "US") %>% 
  filter (!Province_State %in% c("Alaska","Hawaii", "American Samoa",
                  "Puerto Rico","Northern Mariana Islands", 
                  "Virgin Islands", "Recovered", "Guam", "Grand Princess",
                  "District of Columbia", "Diamond Princess")) %>% 
  filter(Lat > 0)
```

```{r}
mybreaks <- c(1, 100, 1000, 10000, 10000)
ggplot(daily_report_9_26_2020_US, aes(x = Long, y = Lat, size = Confirmed)) +
    borders("state", colour = "white", fill = "grey90") +
    geom_point(aes(x=Long, y=Lat, size=Confirmed, color=Confirmed),stroke=F, alpha=0.7) +
    scale_size_continuous(name="Cases", trans="log", range=c(1,7), 
                        breaks=mybreaks, labels = c("1-99",
                        "100-999", "1,000-9,999", "10,000-99,999", "50,000+")) +
    scale_color_viridis_c(option="viridis",name="Cases",
                        trans="log", breaks=mybreaks, labels = c("1-99",
                        "100-999", "1,000-9,999", "10,000-99,999", "50,000+"))  +
  theme_void() + 
    guides( colour = guide_legend()) +
    labs(title = "Anisa Dhana's lagout for COVID-19 Confirmed Cases in the US'") +
    theme(
      legend.position = "bottom",
      text = element_text(color = "#22211d"),
      plot.background = element_rect(fill = "#ffffff", color = NA), 
      panel.background = element_rect(fill = "#ffffff", color = NA), 
      legend.background = element_rect(fill = "#ffffff", color = NA)
    ) +
    coord_fixed(ratio=1.5)
```

#3.
```{r}
report_09_26_2020 <-   read_csv(url("https://raw.githubusercontent.com/CSSEGISandData/COVID-19/master/csse_covid_19_data/csse_covid_19_daily_reports/09-26-2020.csv")) %>% 
  rename(Long = "Long_") %>% 
  unite(Key, Admin2, Province_State, sep = ".") %>% 
  group_by(Key) %>% 
  summarize(Confirmed = sum(Confirmed)) %>% 
  mutate(Key = tolower(Key))

us <- map_data("state")
counties <- map_data("county") %>% 
  unite(Key, subregion, region, sep = ".", remove = FALSE)

state_join_2 <- left_join(counties, report_09_26_2020, by = c("Key"))

ggplot(data = us, mapping = aes(x = long, y = lat, group = group)) + 
  coord_fixed(1.3) + 
  
  borders("state", colour = "black") +
  geom_polygon(data = state_join_2, aes(fill = Confirmed)) +
  scale_fill_gradientn(colors = brewer.pal(n = 5, name = "Purples"),
                       breaks = c(1, 10, 100, 1000, 10000, 100000),
                       trans = "log10", na.value = "White") +
  ggtitle("Number of Confirmed Cases by US County") +
  theme_void() 
```

#4.
```{r}
daily_report_9_26_2020_CO <-   read_csv(url("https://raw.githubusercontent.com/CSSEGISandData/COVID-19/master/csse_covid_19_data/csse_covid_19_daily_reports/09-26-2020.csv")) %>% 
  rename(Long = "Long_") %>% 
  filter(Province_State == "Colorado") %>% 
  group_by(Admin2) %>% 
  summarize(Confirmed = sum(Confirmed)) %>% 
  mutate(Admin2 = tolower(Admin2))
us <- map_data("state")
co_us <- subset(us, region == "colorado")
counties <- map_data("county")
co_county <- subset(counties, region == "colorado")
state_join_2 <- left_join(co_county, daily_report_9_26_2020_CO, by = c("subregion" = "Admin2"))
```

```{r}
library(plotly)
ggplotly(
  ggplot(data = co_county, mapping = aes(x = long, y = lat, group = group)) + 
  coord_fixed(1.3) + 
  geom_polygon(data = state_join_2, aes(fill = Confirmed), color = "white") +
    scale_fill_gradientn(colours = 
                         wes_palette("Moonrise3", 100, type = "continuous")) +
  ggtitle("COVID-19 Cases in CO as of 9/26/2020") +
  labs(x=NULL, y=NULL) +
  theme(panel.border = element_blank()) +
  theme(panel.background = element_blank()) +
  theme(axis.ticks = element_blank()) +
  theme(axis.text = element_blank()) +
    theme(text = element_text(color = "blue"))
)
```

#5.

#Report: COVID19 Cases and Deaths in the State of Colorado as of October 5th 2020

#Confirmed cases by county

```{r, echo = FALSE}
daily_report_10_05_2020_confirmed_CO <- read_csv(url("https://raw.githubusercontent.com/CSSEGISandData/COVID-19/master/csse_covid_19_data/csse_covid_19_daily_reports/10-05-2020.csv")) %>% 
  rename(Long = "Long_") %>% 
  filter(Province_State == "Colorado") %>% 
  filter(Lat != "NA") %>% 
  filter(Long != "NA") %>% 
  group_by(Admin2) %>% 
  summarize(Confirmed = sum(Confirmed)) %>% 
  mutate(Admin2 = tolower(Admin2))
```
```{r, echo = FALSE, optionwarning = FALSE}
us <- map_data("state")
co_us <- subset(us, region == "colorado")
counties <- map_data("county")
co_county <- subset(counties, region == "colorado")
state_join_co <- left_join(co_county, daily_report_10_05_2020_confirmed_CO, by = c("subregion" = "Admin2"))

ggplot(data = co_county, mapping = aes(x = long, y = lat, group = group)) + 
  geom_polygon(data = state_join_co, aes(fill = Confirmed), color = "white") +
    scale_fill_gradientn(colors = brewer.pal(n = 5, name = "OrRd"),
                         trans = "log10") +
  labs(title = "COVID-19 Confirmed Cases in Colorado")
```


#An interactive graph showing the numer of cases in each county:

```{r, echo = FALSE}
library(plotly)
ggplotly(
  ggplot(data = co_county, mapping = aes(x = long, y = lat, group = group)) + 
  coord_fixed(1.3) + 
  geom_polygon(data = state_join_co, aes(fill = Confirmed), color = "white") +
    scale_fill_gradientn(colours = 
                         wes_palette("Royal1", 100, type = "continuous")) +
  ggtitle("COVID-19 Cases in Colorado Counties") +
  labs(x=NULL, y=NULL) +
  theme(panel.border = element_blank()) +
  theme(panel.background = element_blank()) +
  theme(axis.ticks = element_blank()) +
  theme(axis.text = element_blank()) +
    theme(text = element_text(color = "blue"))
)
```

#Confirmed Deaths by County

```{r, echo = FALSE, optionwarning = FALSE}
daily_report_10_05_2020_deaths_CO <- read_csv(url("https://raw.githubusercontent.com/CSSEGISandData/COVID-19/master/csse_covid_19_data/csse_covid_19_daily_reports/10-05-2020.csv")) %>% 
  rename(Long = "Long_") %>% 
  filter(Province_State == "Colorado") %>% 
  filter(Lat != "NA") %>% 
  filter(Long != "NA") %>% 
  group_by(Admin2) %>% 
  summarize(Deaths = sum(Deaths)) %>% 
  mutate(Admin2 = tolower(Admin2))
```
```{r, echo = FALSE, optionwarning = FALSE}
us <- map_data("state")
co_us <- subset(us, region == "colorado")
counties <- map_data("county")
co_county <- subset(counties, region == "colorado")
state_join_co_deaths <- left_join(co_county, daily_report_10_05_2020_deaths_CO, by = c("subregion" = "Admin2"))

ggplot(data = co_county, mapping = aes(x = long, y = lat, group = group)) + 
  geom_polygon(data = state_join_co_deaths, aes(fill = Deaths), color = "white") +
    scale_fill_gradientn(colors = brewer.pal(n = 5, name = "Blues"),
                         trans = "log10") +
  labs(title = "COVID-19 Total Deaths by County: Colorado")
```

#An interactive graph showing the number of deaths in each county:

```{r, echo = FALSE}
library(plotly)
ggplotly(
  ggplot(data = co_county, mapping = aes(x = long, y = lat, group = group)) + 
  coord_fixed(1.3) + 
  geom_polygon(data = state_join_co_deaths, aes(fill = Deaths), color = "white") +
    scale_fill_gradientn(colours = 
                         wes_palette("GrandBudapest2", 100, type = "continuous")) +
  ggtitle("COVID-19 Deaths in Colorado Counties") +
  labs(x=NULL, y=NULL) +
  theme(panel.border = element_blank()) +
  theme(panel.background = element_blank()) +
  theme(axis.ticks = element_blank()) +
  theme(axis.text = element_blank()) +
    theme(text = element_text(color = "blue"))
)
```

Report made in R [@RCoreTeam], using data collected from the Github repository for Novel Coronavirus [@Dong2020].

# References

