---
title: "Lab 9"
author: "Marlee Galloway"
date: "11/24/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## The stringr package

```{r}
library(stringr)
library(tidyverse)
```
## Matching patterns with regular expressions:
# Basic matches

```{r}
x <- c("apple", "banana", "pear")
str_view(x, "an")
```
```{r}
str_view(x, ".a.")
```
```{r}
# To create the regular expression, we need \\
dot <- "\\."

# But the expression itself only contains one:
writeLines(dot)
```
```{r}
# And this tells R to look for an explicit .
x <- c("abc", "a.c", "bef")
str_view(x, "a\\.c")
```
```{r}
x <- "a\\b"
writeLines(x)
```
```{r}
str_view(x, "\\\\")
```

## 14.3.1.1 Exercises
# 1. Explain why each of these strings don’t match a \: "\", "\\", "\\\".

# "/" does not match because it will escape the next character in the R string.
# "//" does not match because it will change back to a / in the regular expression which will then skip the next character in that expression.
# "///" doesn't match because in the regular expression the first two backslashes will become a / which will escape the next character and then the third / will escape the escaped chracter.

## Anchors:

```{r}
x <- c("apple", "banana", "pear")
str_view(x, "^a")
```
```{r}
str_view(x, "a$")
```
```{r}
x <- c("apple pie", "apple", "apple cake")
str_view(x, "apple")
```
```{r}
str_view(x, "^apple$")
```

## 14.3.2.1 Exercises
# 1. How would you match the literal string "$^$"?

# Example of how to match the string: 
```{r}
str_view(c("$^$", "ab$^$sfas"), "^\\$\\^\\$$", match = TRUE)
```

## Character classes and alternatives:

```{r}
# Look for a literal character that normally has special meaning in a regex
x <- c("abc", "a.c", "a*c", "a c")
str_view(x, "a[.]c")
```
```{r}
str_view(x, ".[*]c")
```
```{r}
str_view(x, "a[ ]")
```
```{r}
x <- c("grey", "gray")
str_view(x, "gr(e|a)y")
```

## 14.3.3.1 Exercises
# 1. Create regular expressions to find all words that:
#Start with a vowel:

```{r}
str_subset(stringr::words, "^[aeiou]")
```

#That only contain consonants. (Hint: thinking about matching “not”-vowels.):
```{r}
str_subset(stringr::words, "[aeiou]", negate=TRUE)
```

#End with ed, but not with eed.
```{r}
str_subset(stringr::words, "[^e]ed$")
```

#End with ing or ise.
```{r}
str_subset(stringr::words, "i(ng|se)$")
```

## Repetition:

```{r}
x <- "1888 is the longest year in Roman numerals: MDCCCLXXXVIII"
str_view(x, "CC?")
```
```{r}
str_view(x, "CC+")
```
```{r}
str_view(x, 'C[LX]+')
```
```{r}
str_view(x, "C{2}")
```
```{r}
str_view(x, "C{2,}")
```
```{r}
str_view(x, "C{2,}")
```
```{r}
str_view(x, 'C{2,3}?')
```
```{r}
str_view(x, 'C[LX]+?')
```

## 14.3.4.1 Exercises
# 1. Describe the equivalents of ?, +, * in {m,n} form.

#If the pattern is "?" and {m,n} is {0,1} this means that at most there is one match. The "+" pattern when {m,n} is {1,} means that 1 or more will be matched. The "*" pattern when {m,n} is {0,} means that 0 or more will be matched.

## Grouping and backreferences:

```{r}
str_view(fruit, "(..)\\1", match = TRUE)
```

## 14.3.5.1 Exercises
# 1. Describe, in words, what these expressions will match:
#(.)\1\1: This will match the same character appearing 3 times in a row, ex: "bbb".

#"(.)(.)\\2\\1": This will match a pair of characters followed by the same characters in opposite order, ex: "mggm".

#(..)\1: This will match any pair of characters that is repeated, ex: "b3b3".

#"(.).\\1.\\1": This matches a character followed by a different character followed by the original character followed by another different character and the original again, ex: "mgmjm".

#"(.)(.)(.).*\\3\\2\\1": This matches three characters followed by zero or more characters of any kind followed by the same three characters but in reverse order. E.g. "abcsgasgddsadgsdgcba" or "abccba" or "abc1cba".

## Tools:

# Detect matches

```{r}
x <- c("apple", "banana", "pear")
str_detect(x, "e")
```
```{r}
# How many common words start with t?
sum(str_detect(words, "^t"))
```
```{r}
# What proportion of common words end with a vowel?
mean(str_detect(words, "[aeiou]$"))
```
```{r}
# Find all words containing at least one vowel, and negate
no_vowels_1 <- !str_detect(words, "[aeiou]")
```
```{r}
# Find all words consisting only of consonants (non-vowels)
no_vowels_2 <- str_detect(words, "^[^aeiou]+$")
identical(no_vowels_1, no_vowels_2)
```
```{r}
words[str_detect(words, "x$")]
```
```{r}
str_subset(words, "x$")
```
```{r}
df <- tibble(
  word = words, 
  i = seq_along(word)
)
df %>% 
  filter(str_detect(word, "x$"))
```
```{r}
x <- c("apple", "banana", "pear")
str_count(x, "a")
```
```{r}
# On average, how many vowels per word?
mean(str_count(words, "[aeiou]"))
```
```{r}
df %>% 
  mutate(
    vowels = str_count(word, "[aeiou]"),
    consonants = str_count(word, "[^aeiou]")
  )
```
```{r}
str_count("abababa", "aba")
```
```{r}
str_view_all("abababa", "aba")
```

## 14.4.1.1 Exercises
# 1. For each of the following challenges, try solving it by using both a single regular expression, and a combination of multiple str_detect() calls.
#Find all words that start or end with x.
```{r}
words[str_detect(words, "^x|x$")]
start_with_x <- str_detect(words, "^x")
end_with_x <- str_detect(words, "x$")
words[start_with_x | end_with_x]

```

#Find all words that start with a vowel and end with a consonant.
```{r}
str_subset(words, "^[aeiou].*[^aeiou]$") %>% head()
start_with_vowel <- str_detect(words, "^[aeiou]")
end_with_consonant <- str_detect(words, "[^aeiou]$")
words[start_with_vowel & end_with_consonant] %>% head()
```

#Are there any words that contain at least one of each different vowel? There do not appear to be any.
```{r}
pattern <-
  cross(rerun(5, c("a", "e", "i", "o", "u")),
    .filter = function(...) {
      x <- as.character(unlist(list(...)))
      length(x) != length(unique(x))
    }
  ) %>%
  map_chr(~str_c(unlist(.x), collapse = ".*")) %>%
  str_c(collapse = "|")

str_subset("aseiouds", pattern)

str_subset(words, pattern)

words[str_detect(words, "a") &
  str_detect(words, "e") &
  str_detect(words, "i") &
  str_detect(words, "o") &
  str_detect(words, "u")]
```


## Extract matches:

```{r}
length(sentences)
```
```{r}
head(sentences)
```
```{r}
colours <- c("red", "orange", "yellow", "green", "blue", "purple")
colour_match <- str_c(colours, collapse = "|")
colour_match
```
```{r}
has_colour <- str_subset(sentences, colour_match)
matches <- str_extract(has_colour, colour_match)
head(matches)
```
```{r}
has_colour
```
```{r}
more <- sentences[str_count(sentences, colour_match) > 1]
str_view_all(more, colour_match)
```
```{r}
str_extract(more, colour_match)
```
```{r}
str_extract_all(more, colour_match)
```
```{r}
str_extract_all(more, colour_match, simplify = TRUE)
```
```{r}
x <- c("a", "a b", "a b c")
str_extract_all(x, "[a-z]", simplify = TRUE)

```

## 14.4.2.1 Exercises
# 1. In the previous example, you might have noticed that the regular expression matched “flickered”, which is not a colour. Modify the regex to fix the problem.
```{r}
colour_match2 <- str_c("\\b(", str_c(colours, collapse = "|"), ")\\b")
colour_match2
more2 <- sentences[str_count(sentences, colour_match) > 1]
str_view_all(more2, colour_match2, match = TRUE)
```


## Grouped matches:

```{r}
noun <- "(a|the) ([^ ]+)"

has_noun <- sentences %>%
  str_subset(noun) %>%
  head(10)
has_noun %>% 
  str_extract(noun)
```
```{r}
has_noun %>% 
  str_match(noun)
```
```{r}
tibble(sentence = sentences) %>% 
  tidyr::extract(
    sentence, c("article", "noun"), "(a|the) ([^ ]+)", 
    remove = FALSE
  )
```

## 14.4.3.1 Exercises
# 1. Find all words that come after a “number” like “one”, “two”, “three” etc. Pull out both the number and the word.
```{r}
numword <- "\\b(one|two|three|four|five|six|seven|eight|nine|ten) +(\\w+)"
sentences[str_detect(sentences, numword)] %>%
  str_extract(numword)
```


## Replacing matches:

```{r}
x <- c("apple", "pear", "banana")
str_replace(x, "[aeiou]", "-")
```
```{r}
str_replace_all(x, "[aeiou]", "-")
```
```{r}
x <- c("1 house", "2 cars", "3 people")
str_replace_all(x, c("1" = "one", "2" = "two", "3" = "three"))
```
```{r}
sentences %>% 
  str_replace("([^ ]+) ([^ ]+) ([^ ]+)", "\\1 \\3 \\2") %>% 
  head(5)
```

## 14.4.4.1 Exercises
# 1. Replace all forward slashes in a string with backslashes.
```{r}
str_replace_all("past/present/future", "/", "\\\\")
```


## Splitting:

```{r}
sentences %>%
  head(5) %>% 
  str_split(" ")

```
```{r}
"a|b|c|d" %>% 
  str_split("\\|") %>% 
  .[[1]]
```
```{r}
sentences %>%
  head(5) %>% 
  str_split(" ", simplify = TRUE)
```
```{r}
fields <- c("Name: Hadley", "Country: NZ", "Age: 35")
fields %>% str_split(": ", n = 2, simplify = TRUE)
```
```{r}
x <- "This is a sentence.  This is another sentence."
str_view_all(x, boundary("word"))
```
```{r}
str_split(x, " ")[[1]]
```
```{r}
str_split(x, boundary("word"))[[1]]
```

## 14.4.5.1 Exercises
# 1. Split up a string like “apples, pears, and bananas” into individual components.

```{r}
x <- c("apples, pears, and bananas")
str_split(x, ", +(and +)?")[[1]]
```
