---
title: "Lab9A"
author: "Marlee Galloway"
date: "10/28/2020"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## EXERCISE: Modify examples from Lab 9A

## Conditional statements with if-else

```{r}
x <- runif(1, 0, 10)  
if(x > 5) {
        y <- 10
} else {
        y <- 0
}

```

## for Loops

```{r}
for(m in 5:15) {
        print(m)
}
```

# Using a vector for Loop:

```{r}
x <- c("e", "f", "g", "h")
for(j in 1:4) {
        ## Print out each element of 'x'
        print(x[j])  
}
```

# Use seq_along() to generate an integer sequence based on the length of an object:

```{r}
for(j in seq_along(x)) {   
        print(x[j])
}
```

# Without an index type variable:

```{r}
for(letter in x) {
        print(letter)
}
```

# Curly brackets not needed for one line loops:

```{r}
for(m in 1:4) print(x[m])
```

## Nested for Loops

```{r}
x <- matrix(1:5, 10, 15)
for(m in seq_len(nrow(x))) {
        for(j in seq_len(ncol(x))) {
                print(x[m, j])
        }   
}
```

# While Loops:

```{r}
count <- 0
while(count < 5) {
        print(count)
        count <- count + 1
}
```

# With more than one condition:

```{r}
k <- 2
set.seed(1)
while(k >= 1 && k <= 5) {
        coin <- rbinom(1, 1, 0.5)
        
        if(coin == 1) {  ## random walk
                k <- k + 1
        } else {
                k <- k - 1
        } 
}
print(k)
```

## Functions

# Simple function with no argument:

```{r}
f <- function()
  class(f)
f()
```

# Create a function with a non-trivial function body:

```{r}
f <- function() {
         cat("Amherst, MA\n")
}
f()
```

# Function arguments can be inserted here to determine how many times the argument is printed to the console etc.

```{r}
 f <- function(num) {
         for(i in seq_len(num)) {
                 cat("Amherst, MA\n")
         }}
 f(4)
```

# This funstion returns the total number of characters printed to the console:

```{r}
f <- function(num) {
         hello <- "Amherst, MA\n"
         for(i in seq_len(num)) {
                 cat(hello)
         }
         chars <- nchar(hello) * num
         chars
}
genomics <- f(2)

print(genomics)

```

